Longitudinal Principal Manifold Estimation

### 2023-11-10 paper/main.tex
update figures

\begin{figure}%
  \centering
  \subfloat[\centering Data]{{\includegraphics[width=9cm]{sim_case1_plot2}}}%
  \hfill
  \subfloat[\centering Step 1: Sample Size Reduction]{{\includegraphics[width=9cm]{lpme_data_reduction}}}
  \vfill
  \subfloat[\centering Step 2: Initialization]{{\includegraphics[width=9cm]{lpme_initialization}}}
  \hfill
  \caption{LPME Algorithm Steps}
  \label{fig:lpme_steps}
\end{figure}

@@

The next step in developing the algorithm for estimation of the longitudinal principal manifold is initialization. Different algorithms use different approaches for this step. For example, the PME algorithm uses Isomap (\cite{tenenbaumGlobalGeometricFramework2000}) to develop an initial parameterization, which the algorithm proceeds to iteratively improve. However, implementation of commonly used approaches for initialization in our setting can be problematic. For instance, using ISOMAP to find parameterizations of similarly shaped structures (i.e. the brain region of interest at each time point) is likely to result in inconsistent parameterization directions (i.e. low values in the parameter space may be associated with low values in the embedding space for one run, while low values in the parameter space may be associated with high values in the embedding space in another). Reaching a successful fit in the LPME algorithm requires consistent parameterizations across time points, so a different approach to generating an initial parameterization is needed.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{adni_lhipp}
  \caption{Left Hippocampus, Cognitively Healthy ADNI Participant}
  \label{fig:adni_result}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{adni_lhipp_cross_section}
  \caption{Left Hippocampus Cross Section}
  \label{fig:adni_cross_section}
\end{figure}

>>>>

\begin{figure}%
  \centering
  \subfloat[\centering Data]{{\includegraphics[width=9cm]{sim_case1_plot2}}}%
  \hfill
  \subfloat[\centering Step 1: Sample Size Reduction]{{\includegraphics[width=9cm]{lpme_data_reduction}}}
  \vfill
  \subfloat[\centering Step 2: Initialization]{{\includegraphics[width=9cm]{lpme_initialization}}}
  \hfill
  \subfloat[\centering Steps 3 and 4: Fitting and Tuning]{{\includegraphics[width=9cm]{lpme_fitting}}}
  \caption{LPME Algorithm Steps}
  \label{fig:lpme_steps}
\end{figure}

@@

The next step in developing the algorithm for estimation of the longitudinal principal manifold is initialization. Different algorithms use different approaches for this step. For example, the PME algorithm uses Isomap (\cite{tenenbaumGlobalGeometricFramework2000}) to develop an initial parameterization, which the algorithm proceeds to iteratively improve. However, implementation of commonly used approaches for initialization in our setting can be problematic. For instance, using ISOMAP to find parameterizations of similarly shaped structures (i.e. the brain region of interest at each time point) is likely to result in inconsistent parameterization directions (i.e. low values in the parameter space may be associated with low values in the embedding space for one run, while low values in the parameter space may be associated with high values in the embedding space in another). This is demonstrated in Figure \ref{fig:inconsistent_parameterization}. Reaching a successful fit in the LPME algorithm requires consistent parameterizations across time points, so a different approach to generating an initial parameterization is needed.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{inconsistent_parameterization}
  \caption{Simulated Data with Inconsistent Parameterization}
  \label{fig:inconsistent_parameterization}
\end{figure}

@@

\begin{figure}%
  \centering
  \subfloat[\centering Data]{{\includegraphics[height=22cm]{adni_plots/adni_data_plot}}}%
  \hfill
  \subfloat[\centering LPME Estimate]{{\includegraphics[height=22cm]{adni_plots/adni_lpme_plot}}}
  \hfill
  \subfloat[\centering PME Estimate]{{\includegraphics[height=22cm]{adni_plots/adni_pme_plot}}}

  \caption{Left Hippocampus, Cognitive Healthy ADNI Participant}
  \label{fig:adni_result}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{adni_plots/adni_cross_section}
  \caption{Left Hippocampus Cross Section}
  \label{fig:adni_cross_section}
\end{figure}
